SAT Solving for Vesicle Traffic Systems in Cells

### 2017-07-14 paper/conclusion.tex
many changes

In the experiments, we have demonstrated the power of SMT solvers and the value
of careful encoding of the problems into SMT queries.
%
We manage to scale the tool upto the size that interests biologists.
%
However, there are many further search problems are of interest to
biologists about the VTSs.
%
For example, are there any k-connected graphs that can not be
converted into a VTS?
%
This encoding induces a quantifier alternation in the problem.
%
Therefore, a simple quantifier free encoding will not work.
%
We are planning to use QBF solvers to solve such queries or develop
iterative search algorithm for such queries.
%

>>>>

In the experiments, we have demonstrated the power of SMT solvers and the value
of careful encoding of the problems into SMT queries.
%
We manage to scale the tool upto the size that interests biologists.
%
However, there are many further search problems are of interest to
biologists about the VTSs.
%
For example, are there any k-connected graphs that can not be
annotated into a VTS?
%
This problem induces a quantifier alternation in the problem.
%
Therefore, a simple quantifier free encoding will not work.
%
We are planning to use QBF solvers or develop
iterative search algorithm for such queries.
%

### 2017-07-14 paper/encoding.tex
my changes
many changes

% \subsubsection{Constraints on the structure of the edges, nodes and }
\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints~\eqref{eq:f0} and~\eqref{eq:c3}
to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align}
  \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}\tag{V1}\label{eq:f0}
\end{align}
If a molecule is active on an edge, it should be present on the edge.
%
\begin{align}
  \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\tag{V2}\label{eq:f1}
\end{align}

A molecule should be present to be active on a node.  
\begin{align}
  \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
  \tag{V3}\label{eq:c4}
\end{align}

@@

\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies n_{i,m}\tag{V4}\label{eq:c0}\\
  \bigwedge\limits_{i,j,q,m}  e_{i,j,q,m} \limplies n_{j,m}\tag{V5}\label{eq:c1}
\end{align}

%C1 =  e_ijk -> aik and e_ijk -> ajk

Self edges are not allowed.

\begin{align}
   \bigwedge\limits_{i,q} \neg e_{i,i,q}\tag{V6}\label{eq:c2}
\end{align}

Since a molecule cannot be both SNARE and tether,
we set diagonal blocks in pairing matrix to be 0's.
It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
\begin{align}
  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
  \tag{V7}\label{eq:c3}
\end{align}

\subsubsection{Well-fused constraints}
Boolean constraint~\eqref{eq:fuse1} encodes that each edge must fuse with
its destination node.
%
Boolean constraint~\eqref{eq:fuse2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align}
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies \bigvee_{m,m^{\prime}} (b_{i,j,q,m} \land a_{j,m^{\prime}} \land p_{m,m^{\prime}})
  \tag{V8}\label{eq:fuse1}  \\
\bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies \neg \bigvee_{j \neq j^{\prime}, m^{\prime\prime}} ( a_{j^{\prime},m^{\prime\prime}} \land p_{m,m^{\prime\prime}})
  \tag{V9}\label{eq:fuse2}  
\end{align}

@@

\subsubsection{Regulation on nodes and edges}
We are considering several models that differ in constraints on
the activity of molecules.
%
We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
the varying constraints.
%
The following formula encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} = a_{i,m}    \tag{Ann}\label{eq:ann}
\end{align}
The following formula encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\tag{Anb}\label{eq:anb}
\end{align}
The following formula encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\tag{Aen}\label{eq:aen}
\end{align}
The following formula encodes that activity of a molecule $m$ on the edge is
defined by a Boolean function $\edgef_m$ of presence of molecules present on that edge.
\begin{align}
   \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
  \tag{Aeb}\label{eq:aeb}
\end{align}
%
The following formula encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align}
   \bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \limplies  \bigwedge_{m' \neq m} (p_{m,m'} \limplies e_{i,j,q,m'})] \lequiv (\neg b_{i,j,q,m} \land  \bigwedge_{m' \neq m, p_{m,m'}} \neg b_{i,j,q,m'})
  \tag{Aep}\label{eq:aep}
\end{align}
% \[ \bigwedge\limits_{i,j,q, k}  (e_{i,j,q,k} \land [ ( \bigvee_{k'} p_{k,k'} \land \bigwedge_{k' != k} p_{k,k'} \limplies e_{i,j,q,k'})] ) \limplies \neg b_{i,j,q,k}) \, \]

% \[ \bigwedge\limits_{i,j, q, k}  (e_{i,j,q,k} \land \neg [ ( \bigvee_{k'} p_{k,k'} \land \bigwedge_{k' != k} p_{k,k'} \limplies e_{i,j,q,k'})] ) \limplies b_{i,j,q,k}) \, \]

% - F0: b(i,j,k) $\limplies$ e(i,j,k)

\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following formula recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align}
  \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
  \tag{R1}\label{eq:reach1}
\end{align}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.\ashu{disjunction needed??}
\begin{align}
 \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies (r_{j,i,m,0}  \lor r_{j,i,m,1} \lor ... \lor r_{j,i,m,p})  
  \tag{R2}\label{eq:reach2}
\end{align}

@@

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
The following formulas encode the relevant constraints for reachability in the modified VTS.

The following formulas encode that only existing edges can be dropped
and exactly $k-1$ edges are dropped.
\begin{align}
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \tag{D1}\label{eq:drop1}\\
  \sum_{i,j,q} d_{i,j,q} = k-1
  \tag{D2}\label{eq:drop2}
\end{align}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following formula, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align}
\bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}  
  \tag{D3}\label{eq:drop3}
\end{align}
Since we are looking for a graph that is not $k$-connected,
the following formula encodes that there are nodes $i$ and $j$ such that
there is no path between them in the underlying undirected graph.
\begin{align}
   \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
  \tag{D4}\label{eq:drop4}
\end{align}

@@

We have encoded several versions of the models of vesicle traffic systems listed. Starting from most basic version (A) where every present molecule is considered to be active. Other cases includes different combinations of constraints on the activity on nodes and edges. Most general restriction is activity of the molecule is dependent on the other molecules presence on the node (a Boolean function of N-1 parameters) (B,C,D,F,G,H). In very special case (E,F,G,H) activity of the molecules on the edges is dictated by pairing matrix, we call this process ``pairing inhibition". 
%In this case if the corresponding pairing molecules present on the edge with the molecules , both the molecule and all corresponding molecules gets inhibited/become inactive.

@@

\begin{enumerate}[label=\Alph*]
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
\item Version F with additional constraint that every edge is distinct.
%\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition with every edge is distinct.
%\item Activity of molecules on the nodes is based on Boolean function of presence of other moleculess and on edge by pairing inhibition with self-edges are allowed.
%\item Version F but self-edges are allowed.
\end{enumerate}

We have applied the tool for the above variations for searching the vesicle traffic systems that satisfies
their corresponding properties.

\input{pm-1-matrix}
\input{v2-dia}

After creating the formula using different Boolean variables we ship them directly to the solver. In case of a satisfying (SAT) assignment of the added constraints the Z3 theorem prover gives a model with the variable assignment. In case no possible satisfying assignment the solver returns UNSAT.
%
%

>>>>

% \subsubsection{Constraints on the structure of the edges, nodes and }
\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints~\eqref{eq:f0} and~\eqref{eq:c3}
to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align}
  \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}\tag{V1}\label{eq:f0}
\end{align}
If a molecule is active on an edge, it should be present on the edge.
%
\begin{align}
  \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\tag{V2}\label{eq:f1}
\end{align}
A molecule should be present to be active on a node.  
\begin{align}
  \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
  \tag{V3}\label{eq:c4}
\end{align}

@@

\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )\tag{V4}\label{eq:c0}
\end{align}
Self edges are not allowed. 
\begin{align}
   \bigwedge\limits_{i,q} \neg e_{i,i,q}\tag{V5}\label{eq:c2}
\end{align}
Since a molecule cannot be both SNARE and tether,
we set diagonal blocks in pairing matrix to be 0's.
It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
\begin{align}
  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
  \tag{V6}\label{eq:c3}
\end{align}

\subsubsection{Well-fused constraints}
Boolean constraint~\eqref{eq:fuse1} encodes that each edge must fuse with
its destination node.
%
Boolean constraint~\eqref{eq:fuse2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align}
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies \bigvee_{m,m^{\prime}} (b_{i,j,q,m} \land a_{j,m^{\prime}} \land p_{m,m^{\prime}})
  \tag{V7}\label{eq:fuse1}  \\
\bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies \neg \bigvee_{j \neq j^{\prime}, m^{\prime\prime}} ( a_{j^{\prime},m^{\prime\prime}} \land p_{m,m^{\prime\prime}})
  \tag{V8}\label{eq:fuse2}  
\end{align}

@@

\subsubsection{Regulation on nodes and edges}
We are considering several models that differ in constraints on
the activity of molecules.
%
We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
the varying constraints.
%
The following formula encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} = a_{i,m}    \tag{Ann}\label{eq:ann}
\end{align}
The following formula encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\tag{Anb}\label{eq:anb}
\end{align}
The following formula encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\tag{Aen}\label{eq:aen}
\end{align}
The following formula encodes that activity of a molecule $m$ on the edge is
defined by a Boolean function $\edgef_m$ of presence of molecules present on that edge.
\begin{align}
   \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
  \tag{Aeb}\label{eq:aeb}
\end{align}
%
The following formula encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align}
   \bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \limplies  \bigwedge_{m' \neq m} (p_{m,m'} \limplies e_{i,j,q,m'})] \lequiv (\neg b_{i,j,q,m} \land\;\;\;\;  \bigwedge_{\mathclap{m' \neq m, p_{m,m'}}} \neg b_{i,j,q,m'})
  \tag{Aep}\label{eq:aep}
\end{align}
%
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following formula recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align}
  \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
  \tag{R1}\label{eq:reach1}
\end{align}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align}
 \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,|N|}
  \tag{R2}\label{eq:reach2}
\end{align}

@@

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
%
Boolean constraints~\ref{eq:drop1}-\ref{eq:drop4}
encode the relevant constraints for reachability
in the modified VTS. 
%
The following formulas encode that only
existing edges can be dropped and exactly $k-1$ edges are dropped.
\begin{align}
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \tag{D1}\label{eq:drop1}\\
  \sum_{i,j,q} d_{i,j,q} = k-1
  \tag{D2}\label{eq:drop2}
\end{align}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following formula, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align}
\bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}  
  \tag{D3}\label{eq:drop3}
\end{align}
Since we are looking for a graph that is not $k$-connected,
the following formula encodes that there are nodes $i$ and $j$ such that
there is no path between them in the underlying undirected graph.
\begin{align}
   \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
  \tag{D4}\label{eq:drop4}
\end{align}

@@

We have encoded several versions of the models of VTSs listed.
%
The cases includes different combinations of constraints on the
activity on nodes and edges.
%

%In this case if the corresponding pairing molecules present on the edge with the molecules , both the molecule and all corresponding molecules gets inhibited/become inactive.

@@

\begin{enumerate}[label=\Alph*]
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
\item Version F with additional constraint that every edge is distinct.\ashu{encode this??}
%\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition with every edge is distinct.
%\item Activity of molecules on the nodes is based on Boolean function of presence of other moleculess and on edge by pairing inhibition with self-edges are allowed.
%\item Version F but self-edges are allowed.
\end{enumerate}
%
The version (A) is most basic where every present molecule is active.
%
In versions (B-D,F-H),
most general restriction is activity of the molecule is dependent on
the other molecules presence on the node.
%
In the versions (E-H), activity of the molecules on the edges
is dictated by pairing matrix; we call this process ``pairing
inhibition".
%
In the column 2-3 of table~\ref{table1}, we list the constraints involved
in each version.

% We have applied the tool for the above variations for searching the
% VTSs that satisfies their corresponding properties. 
%
% After creating the formula using different Boolean variables 
% we ship them directly to the solver.
% %
% In case of a satisfying (SAT) assignment of the added constraints the
% Z3 theorem prover gives a model with the variable assignment.
% %
% In case no possible satisfying assignment the solver returns UNSAT.

%
%

### 2017-07-14 paper/experiments.tex
many changes

Given a system that follows the VTS, we are interested in its connectivity (graph) property. We have implemented the encoding of the constraints in the python interface of $\zthree$. The basic idea is to constraint the variables such that vesicular transport rules are respected and using constraints (D0-D5) reason about the connectivity of the underlying structure (graph).

@@

%Besides using the default Z3 solvers we have used for these experiments \ankit{Z3, picoSAT, Lingeling}. The performance of Z3 was ...
%
% \Ankit{In other solver variations ...}
Our tool allows the user to choose a model and the size of the network besides other parameters like connectivity and number of parallel edges. It uses Z3 Python interface to build the needed constraints and applies $\zthree$ solver on the constraints to find a model ( a satisfying assignment that respects the constraints). This tool also translates the satisfying model found by $\zthree$ into a vesicle traffic system and presents a visual output to the user in form of annotated graph. The output graph will satisfy the underlying rules of the system, so it always is a valid vesicle transport network. The resultant graph provides an overview of the underlying vesicle transport system where directed labeled edges and nodes provide the complete information about the system. Dashed lines are used in the resultant visual network output to specify the dropped edges to make the graph disconnected, which gives information about the connectivity of the graph.

\input{tbl-qf-graph}
%\input{tbl-qf1-graph}
In Table 1. 
%~\ref{tab:qf-graph},
we present the running times for the search of vesicle traffic systems of sizes 2 to 10 that satisfies the model variants. For example, the formula for variation F, the total number of compartments (N) equals to 10, returns in 129.78 minutes (7786.8 secs) with a SAT result. In comparison, CBMC encoding results in OUT OF MEMORY for the same variation and same compartment size. Hence with the use of this novel encoding, we are able to scale the system to a much larger compartmentalized system, especially to eukaryotic cells with a total number of ten compartments. This additional cushion of scalability can provide us leverage to ask for related open questions that were not previously possible, for example, what is the minimum number of molecules required to satisfy the vesicular traffic system in eukaryotes.

\input{cz-comparision}

% REQUIRE COMPARISION WITH CBMC RUN TIME ?
The experiments were done on a machine with Intel(R) Core(TM) i3-4030U CPU @ 1.90GHz processor and 4GB RAM. On this machine, the CBMC encoding ran out of memory (M/O) for N $\geq$ 6 whereas new novel encoding was able to handle the case fairly easily. We think the main reasons for such an improvement are the following:
\begin{enumerate}
\item Direct encoding into the solver rather than using CFE (CBMC C front end), which treats front end variables as C variable and similarly other C overheads.
\item Current novel encoding is fine-tuned to the structure of the problem.
\item Improved encoding based on reachability rather than complete enumeration and non-determinism.
\end{enumerate} 
% 
$\zthree$ was able to solve the constraints up to 14-18 nodes (based on the versions).
%

@@

\begin{table}[!ht]
\centering
\def\arraystretch{1.6}
\caption{
{\bf Activity regulation of molecules and corresponding connectivity of the graph.}}
  \begin{tabular}{|c|c|c|c|}
    \hline
  {\multirow{2}{*}{\textbf{Version}}}  & {\multirow{2}{*}{\textbf{Constraints}}} &  \multicolumn{2}{c|}{\textbf{Graph connectivity}}  \\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
   \cline{3-4}
   {} & {} & \textbf{\textbf{Least connected}} & \textbf{Sampling: Guarantee}\\
    %\hhline{~--}
    \hline

%    A. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nn A\_en & No graph  \\ \hline
%B. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_en & No graph  \\ \hline
%C. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nn A\_eb & 3-connected  \\  \hline
%D. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nn A\_ep & No graph  \\ \hline
%E. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_eb & 2-connected  \\ \hline
%F. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_ep & 4-connected  \\ \hline
%G. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_eb C\_ed & 3-connected \\ \hline

A. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nn A\_en & No graph & No graph \\ \hline
B. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_en & No graph & No graph \\ \hline
C. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nn A\_eb & 3-connected & 3-connected \\  \hline
D. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nn A\_ep & No graph & No graph \\ \hline
E. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_eb & 2-connected & 3-connected \\ \hline
F. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_ep & 4-connected & 4-connected \\ \hline
G. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb A\_eb C\_ed & 3-connected & 3-connected \\ \hline
%H. & V1-V7 V8 V9 R1 R2 D1-D4 A\_nb  A\_eb + C\_es & No-idea & No-idea \\ \hline

% C_es :Self edges are allowed
% C_ed : Every edge is distinct 
  \end{tabular}
\label{table1}
\end{table}

>>>>

\begin{table}[t]
\centering
% \def\arraystretch{1.6}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    {\multirow{2}{*}{\textbf{Version}}}  & 
    \multicolumn{2}{c|}{\multirow{2}{*}{\textbf{Constraints}}} & 
    \multicolumn{2}{c|}{\textbf{Graph connectivity}}  \\
   \cline{4-5}
   &  &  & \textbf{\textbf{Least connected}} & \textbf{Sampling: Guarantee}\\ \hline

A. & \multirow{5}{*}{
     \makecell{ \ref{eq:f0}--\ref{eq:fuse2},\\
     \ref{eq:reach1},\ref{eq:reach2},\\
     \ref{eq:drop1}--\ref{eq:drop4}}
     } 
& \ref{eq:ann},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
B. & & \ref{eq:anb},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected & 3-connected \\  \cline{3-5}
D. & & \ref{eq:ann},\ref{eq:aep} & No graph & No graph \\ \cline{3-5}
E. & & \ref{eq:anb},\ref{eq:aeb} & 2-connected & 3-connected \\ \cline{3-5}
F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected & 4-connected \\ \cline{3-5}
G. & & \ref{eq:anb},\ref{eq:aeb},\ref{eq:ced} & 3-connected & 3-connected \\ \hline

% C_es :Self edges are allowed
% C_ed : Every edge is distinct 
  \end{tabular}
\caption{{\bf Activity regulation of molecules vs. graph connectivity.}}
\label{tab:grph}
\end{table}

% Given a system that follows the VTS, we are interested in its connectivity (graph) property.

% The basic idea is to constraint the variables such that vesicular
% transport rules are respected and using constraints (D0-D5) reason
% about the connectivity of the underlying structure (graph).

@@

%Besides using the default Z3 solvers we have used for these experiments \ankit{Z3, picoSAT, Lingeling}. The performance of Z3 was ...
%
% \Ankit{In other solver variations ...}

We have implemented the encoding of the constraints for each version
using the python interface of $\zthree$ in a tool.
%
Our tool allows the user to choose a model and the size of the network
besides other parameters like connectivity and number of parallel
edges.
%
It uses Z3 Python interface to build the needed constraints and
applies $\zthree$ solver on the constraints to find a model ( a
satisfying assignment that respects the constraints).
%
This tool also translates the satisfying model found by $\zthree$ into
a VTS and presents a visual output to the user in form of annotated graph.
%
The output graph will satisfy the underlying rules of the system, so
it always is a valid vesicle transport network.
%
The resultant graph provides an overview of the underlying VTS
where directed labeled edges and nodes provide the
complete information about the system.
%
We also visually report the dropped edges to make the graph
disconnected, which gives information about the connectivity of the
graph.

% Dashed lines are used in the resultant visual network output to
% specify the dropped edges to make the graph disconnected, which gives
% information about the connectivity of the graph.
\ashu{discuss table~\ref{tab:grph}}

\input{tbl-qf-graph}

In Table~\ref{tab:qf-grabh}, we present the running times for the search of
vesicle traffic systems of sizes 2 to 10 that satisfies the model
variants.
%
The experiments were done on a machine with Intel(R) Core(TM) i3-4030U
CPU @ 1.90GHz processor and 4GB RAM.
%
We have compared our performance with the performance of our earlier 
CBMC based implementation.
%
For example, the formula for variation F, the total number of
compartments ($|N|$) equals to 10, returns in 129.78 minutes (7786.8 secs)
with a SAT result.
%
In comparison, CBMC results in OUT OF MEMORY for $|N|$ greater than 5.
%
Hence with the use of this novel encoding, we are able to scale the
system to a much larger compartmentalized system, especially to
eukaryotic cells with a total number of ten compartments.
%
This additional scalability can provide us leverage to ask for related
open questions that were not previously possible, for example, what is
the minimum number of molecules required to satisfy the vesicular
traffic system in eukaryotes.
%
Furthermore, we experimented with limits of our tools and found
that $\zthree$ was able to solve the constraints up to $\sim{14-18}$ nodes.

% \input{cz-comparision}

% REQUIRE COMPARISION WITH CBMC RUN TIME ?

% %
% On this machine, the CBMC encoding ran out of memory (M/O) for N
% $\geq$ 6 whereas new novel encoding was able to handle the case fairly
% easily.
% %
% We think the main reasons for such an improvement are the following:
% \begin{enumerate}
% \item Direct encoding into the solver rather than using CFE (CBMC C front end), which treats front end variables as C variable and similarly other C overheads.
% \item Current novel encoding is fine-tuned to the structure of the problem.
% \item Improved encoding based on reachability rather than complete enumeration and non-determinism.
% \end{enumerate} 
% % 
% %

@@
